Actuator line modeling of vertical-axis turbines

### 2016-04-29 paper.tex
Add two introductory paragraphs; resolves #11

Despite its apparent---but not completely certain---effectiveness for predicting
the performance and wake of a single turbine, blade-resolved CFD presents a huge
computational expense since it must resolve fine details of the blade boundary
layers, and must be performed in three dimensions, which will preclude its use
for array analysis until the availability of computing power increases
sufficiently. It is therefore necessary to explore simpler models that can
predict the turbine loading and flow field with acceptable fidelity, but that
are economical enough to not require high performance computing, at least for
individual devices.

Very fast solution times can be obtained with blade element momentum (BEM)
models, such as the double multiple streamtube (DMST) approached developed by
Paraschivoiu~\cite{Para1988}. However, their poor performance for high solidity
(or chord-to-radius ratio $c/R$) turbines \cite{Joo2015}, makes them less
attractive, particular for marine hydrokinetic applications. Their description
of the flow field is very crude as well, simply computing the momemtum contained
within discretized streamtubes, which cannot resolve any of the complicated flow
structures shed by the rotor.

>>>>

Vertical-axis (cross-flow) turbines (VATs or CFTs) were the subject of
significant research and development in the 1970s through the 1990s by groups
like Sandia National Labs in the US~\cite{Sutherland2012} and the National
Research Council of Canada~\cite{Para2002}. Despite minor commercial success for
large scale onshore wind applications, vertical-axis turbines were virtually
abandoned in favor of horizontal-axis (or axial-flow) turbines, as they
generally are more efficient and don't encounter the high levels of fatigue
loading that VATs do. Today, however, there is renewed interest in VATs for
marine hydrokinetic (MHK) applications~\cite{ORPC2012}, urban
wind~\cite{Awan2013}, and smaller scale, tightly spaced wind
farms~\cite{Dabiri2011, Kinzel2012}, thanks to their relatively faster wake
recovery.

The mean near-wake structure of vertical-axis turbine has been shown to be
largely dominated by the effects of tip vortex shedding, which induces levels of
recovery due to vertical advection significantly larger than those from
turbulent fluctuations~\cite{Bachant2015-JoT}---an effect not seen in HAT wakes.
As interest shifts to designing and analyzing arrays of VATs, it is necessary to
determine the effectiveness of various numerical modeling techniques to
replicate VAT near-wake dynamics, such that wake recovery is accurately
predicted, leading to accurate assessment of optimal array spacing.

Reynolds-averaged Navier--Stokes (RANS) turbulence models computed on 3-D
body-fitted (blade-resolved) grids can do a good job predicting the mean
performance and near-wake structure of a VAT, though their effectiveness depends
on the turbulence model applied~\cite{Bachant2016-BR-CFD}. However, 3-D
blade-resolved RANS presents a huge computational expense---on the order of
1,000 CPU hours per simulated second with contemporary hardware---since it must
resolve fine details of the blade boundary layers, which will preclude its use
for array analysis until the availability of computing power increases
sufficiently. It is therefore necessary to explore simpler models that can
predict the turbine loading and flow field with acceptable fidelity, but that
are economical enough to not require high performance computing, at least for
individual devices.

Very fast solution times can be obtained with blade element momentum (BEM)
models, such as the double multiple streamtube (DMST) approached developed by
Paraschivoiu~\cite{Para1988}. However, their poor performance for high solidity
(or chord-to-radius ratio $c/R$) turbines \cite{Joo2015}, makes them less
attractive, particular for marine hydrokinetic applications, where rotors of the
same size will encounter approximately an order of magnitude higher torque
compared to wind turbines. Their description of the flow field is very crude as
well, simply computing the momemtum contained within discretized streamtubes,
which cannot resolve any of the complicated flow structures shed by the rotor.
